Growing Up with Nell: A Narrative Interface for Literacy

### 2012-03-12 paper.tex
Spell-checking.
Rewrite Personalization section to lead with the Multiple Intelligences stuff.

%As depicted in Figure~\ref{fig:nell},
Nell's characters are
always-available agents layered above a particular system \textit{activity}
(application) which provides specialized functionality.  The agents
are not always foregrounded: constructionist learning occurs when the
child plays freely to ``make things'' with the base activity.
The handwriting tutor is fundamentally a drawing activity; the
adventure involving the magical musical lock is also a music-making activity.
%  ``Making things'' is the key to constructionist learning.
The narrative system is
hooked into each activity to provide passive guidance
(congratulating the child when it notices they've drawn a letter),
active guidance (Apple-Guide--style~\cite{powers:appleguide}
contextual help), or system services (switching activities; jumping into a
related story module).

@@

%Children learn in different ways...

Children often decorate their belongings, and the
child owning a Nell system is encouraged to personalize it.
Superficially, colors can be changed and characters can be renamed and
dressed up.  But Nell also supports deeper story personalization.

\textit{Avatar Services} stores all choices made by the child and make them
accessible to story modules and activities.  Accomplishments
(completed lessons) are also stored.  Nell can thus provide multiple
alternative story modules for a given plot point/lesson, and select
between them based on either the explicit choices made by the child
(``fractions in outer space,'' if the child previously indicated a
preference for space-themed lessons) or prior success with a given
lesson style (a large number of accomplishments in musical-rhythmic
tasks suggests a rhythmic approach to fractions).  In this manner we
can engage the child's interests and cater to different learning
styles.%~\cite{gardner:mi}.%%%%%%% <- lead with this.

Moving further, this record of past choices and accomplishments can be
reviewed.  The narrative history is stored in the \textit{Journal},
where it can form the core of a Portfolio~\cite{stefanakis:portfolios}.
In order to encourage fearless play and experimentation, the Journal
also supports pervasive undo; you can replay the narrative starting at
any past point in the journal.

In the future, we expect to allow further narrative personalization
via recombination of story elements.
Lebowitz~\cite{lebowitz:universe85} and Riedl~\cite{riedl:planning}
show how a planner can be used to recombine and adapt story fragments.
We have less ambition than the cited work: instead of
attempting to generate thousands of stories from tens of templates, we
expect to select and then modestly adapt from hundreds of story
modules created in a decentralized manner by teachers---and eventually
by the students themselves.

@@

Children grow. To accommodate them, Nell provides a ``low floor and
no ceiling.''  The low
floor is a very simple base system, usable with no literacy skills or
external guidance.  The serial progression of story modules grows with
the child to tackle topics of increasing difficulty.
Later, the ability for third
parties to independently author story modules ensures that
Nell can continue to grow.  Finally, there are no external dependencies
limiting a child's ability to participate in this process: a Nell
system contains a story editor and everything necessary to
author and publish story modules.

@@

%%  LocalWords:  personalization Lebowitz

>>>>

%As depicted in Figure~\ref{fig:nell},
Nell's characters are
always-available agents layered above a particular system \textit{activity}
(application) which provides specialized functionality.  The agents
are not always foregrounded: constructionist learning occurs when the
child plays freely to ``make things'' with the base activity.
The handwriting tutor is fundamentally a drawing activity; the
adventure involving the magical musical lock is also a music-making activity.
%  ``Making things'' is the key to constructionist learning.
The narrative system is
hooked into each activity to provide passive guidance
(congratulating the child when it notices they've drawn a letter),
active guidance (Apple-Guide--style~\cite{powers:appleguide}
contextual help), or system services (switching activities; jumping into a
related story module).  A system-wide \textit{achievement} system provides
goals and rewards.

@@

Children learn in different ways.  Nell provides multiple alternative
story modules for its plot points/lessons to engage the child's
interests and cater to multiple intelligences~\cite{gardner:mi}.
Selecting between the alternatives can depend on either explicit
choices made by the child (``fractions in outer space,'' if the child
chose a space-themed story or previously indicated an interest in space)
or prior success with a given
lesson style (a large number of accomplishments in musical-rhythmic
tasks suggests taking a rhythmic approach to fractions).  Similarly, several
different achievements can coexist, rewarding different ways of
accomplishing the same pedagogical goal.

This record of past choices and accomplishments is stored in a
\textit{Journal} and can be reviewed by the child.  The contents of the
Journal can be rearranged to create a Portfolio~\cite{stefanakis:portfolios}
demonstrating the child's progress.
In order to encourage fearless play and experimentation, the Journal
also supports pervasive undo; the child can rewind and replay the
narrative starting at any past point in the journal.

Content can also be remixed for further personalization.
The child is encouraged to rename characters and change clothing,
colors, and other superficial details.  In the future
we expect to allow further narrative customization by
recombination of story elements.
Lebowitz~\cite{lebowitz:universe85} and Riedl~\cite{riedl:planning}
show how a planner can be used to recombine and adapt story fragments.
We have less ambition than the cited work: instead of
attempting to generate thousands of stories from tens of templates, we
hope to select and then modestly adapt from hundreds of story
modules created in a decentralized manner by teachers---and eventually
by the students themselves.

@@

% CSA: three axes of growth
%  1. serial tech tree of story material; more advanced topics
%  2. turtlescript -- metacircular understanding of the system itself
%  3. community -- no limit to content creation, including by the child.

Children grow. To accommodate them, Nell provides a ``low floor and
no ceiling.''  A very simple base system provides the low
floor, usable with no literacy skills or
external guidance.  The serial progression of story modules grows with
the child to tackle topics of increasing difficulty.
Later, the ability for third
parties to independently author story modules ensures that
Nell can continue to grow.  Finally, there are no external dependencies
limiting a child's ability to participate in this process: a Nell
system contains a story editor and everything necessary to
author and publish story modules.

@@

%%  LocalWords:  personalization Lebowitz TurtleScript AIML Loebner
%%  LocalWords:  AliceBot modularity Rahel Ananian Akamai Papert JSON
%%  LocalWords:  Riedl WebGL touchpad Curveship Rete TinkRBook
